---
title: 'Peer-graded Assignment: Statistical Inference Course Project (Part-1)'
author: "Sanjaya J Shetty"
date: "24/11/2020"
output:
  html_document: default
  word_document: default
  pdf_document: default
---

# Simulation Exercise

#### In the first part we look into the distribution of the sample mean of value simulated from an exponential distribution with λ=0.2 with the Central Limit Theorem. The sample mean will be taken over 40 values and repeated for 1000 simulations.

## 1) Show the sample mean and compare it to the theoretical mean of the distribution.

```{r}
set.seed(2020)
n <-  40
Sim <- 1000
Lambda <- 0.2
```

### Sample mean
```{r}
SampleMean <- NULL

for(i in 1:Sim){
  SampleMean <- c(SampleMean, mean(rexp(n, Lambda)))
}


mean(SampleMean)
```

### Theoretical mean

```{r}
theoretical_mean <- 1/Lambda
theoretical_mean
```

## 2) Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution

### Sample variance
```{r}
SampleVar <- var(SampleMean)
SampleVar
```

### Theoretical Variance
```{r}
theoretical_var <- ((1/Lambda)*(1/sqrt(n)))^2
theoretical_var
```

## 3) Show that the distribution is approximately normal.

### Histogram
```{r}
hist(SampleMean, col = "Yellow",
     xlab = "Means", probability = T, breaks = n,main = "Normal Distribution Comparsion")
x <- seq(min(SampleMean), max(SampleMean), length = 200)
lines(x, dnorm(x, mean = 1/Lambda, sd = (1/Lambda/sqrt(n))), pch = 25, col = "red")
```

### Q-Q Plot
```{r}
qqnorm(SampleMean)
qqline(SampleMean, col = "orange")
```
