---
title: 'Peer-graded Assignment: Statistical Inference Course Project (Part 2)'
author: "Sanjaya J Shetty"
date: "24/11/2020"
output: html_document
---

# Basic inferential data analysis.

## Load the libraries
```{r, warning=FALSE, message=FALSE}
library(ggplot2)
library(datasets)
```

### 1. Load the ToothGrowth data and perform some basic exploratory data analyses
```{r}
data("ToothGrowth")

str(ToothGrowth)
head(ToothGrowth)
tail(ToothGrowth)
```

### 2) Provide a basic summary of the data.
```{r}
summary(ToothGrowth)

ggplot(ToothGrowth, aes(x=supp, y=len))+ 
        geom_boxplot(aes(fill=supp))+
        labs(x = 'Type of Supplement', y = "Length of the Tooth", title = "Relationship between Tooth length and Supplement")+
        guides(fill=guide_legend(title="Supplement"))

```

```{r}
ggplot(aes(x = factor(dose), y = len), data = ToothGrowth) + 
    geom_boxplot(aes(fill = factor(dose)))+
    labs(x = "Type of Dose", y = "Length of the Tooth" ,title="Relationship between Tooth Length and Dose")+
    guides(fill=guide_legend(title="Dose"))
```

### 3) Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose.
```{r}
t.test(len~supp, data = ToothGrowth, paired = F)
```

```{r}
t.test(len~supp, data = subset(ToothGrowth, dose ==0.5 ), paired = F)
```

```{r}
t.test(len~supp, data = subset(ToothGrowth, dose ==1 ), paired = F)
```

```{r}
t.test(len~supp, data = subset(ToothGrowth, dose ==2 ), paired = F)
```


# Conclusions

### By looking into p-value (with Alpa = 5%) from the above Hypothesis testing we could conclude that

 + The supplement OJ and VJ doesn't have a significant      impact on the average tooth growth
 + The dose 0.5 and 1 does have any impact on the           average tooth growth
 + The dose 2 doesn't have a significant impact on the      average tooth growth
