---
title: 'Peer-graded Assignment: Prediction Assignment Writeup'
author: "Sanjaya J Shetty"
date: "25/11/2020"
output: html_document
---


# Load all the library

```{r, echo=TRUE, warning=FALSE, message=FALSE}
set.seed(1234)

library(ggplot2)
library(readr)
library(dplyr)
library(caret)
library(rpart)
library(rpart.plot)
library(rattle)
library(randomForest)
```

# loading and cleaning the dataset

```{r, results='hide'}
training_url <- ('https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv')
download.file(training_url, destfile = "training_data.csv")

testing_url <-  ('https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv')
download.file(testing_url, destfile = "testing_data.csv")

train_data <- read.csv('training_data.csv', header = T, sep = ',', na.strings = c("", "#DIV/0!", "NA"))
test_data <- read.csv('testing_data.csv', header = T, sep = ',', na.strings = c("", "#DIV/0!", "NA"))
```


### Removing irrevelvent column entries (i.e, near-zero variance predictors, the first 5 columns seems not effect the classe variable and NA columns)

### Removing NA columns
```{r}
train_data <- train_data[,which(colSums(is.na(train_data)) == 0)]
test_data <-  test_data[,which(colSums(is.na(test_data)) == 0)]
```

### Remove near zero variance predictors
```{r}
train_data <-  train_data[, -(nearZeroVar(train_data))]
test_data <-  test_data[, -(nearZeroVar(test_data))]
```

### Remove first 5 columns
```{r}
train_data <- train_data[,-(1:5)]
test_data <-  test_data[, -(1:5)]

dim(train_data)
dim(test_data)
```

### Create a training and validation sets from the train dataset
```{r}
inTrain <-  createDataPartition(train_data$classe, p =0.7, list = FALSE )

training <- train_data[inTrain,]
validation <- train_data[-inTrain,]

training$classe <- as.factor(training$classe)
validation$classe <- as.factor(validation$classe)
```

## Fitting Model

### With Random Forest
```{r}
Model_1 <- train(classe~., method = "rf", data = training, trControl = trainControl(method="cv"), number = 5, prox = T, list =F, ntree = 100 )
Pred_1 <- predict(Model_1, validation)

confusionMatrix(Pred_1, validation$classe)
```


### With Liner Discriminant Analysis
```{r}
Model_2 <- train(classe~., data = training, method = "lda")
Pred_2 <- predict(Model_2, validation)

confusionMatrix(Pred_2, validation$classe)
```

### Since the accuracy of random forest shows nearly 99.7%, So we would be using that model to predict the test data using Model_1

## Final Prediction for the test data
```{r}
predict(Model_1, test_data)
```

