---
title: 'Peer-graded Assignment: Regression Models Course Project'
author: "Sanjaya J Shetty"
date: "25/11/2020"
output: html_document
---

# Instruction
#### You work for Motor Trend, a magazine about the      automobile industry. Looking at a data set of a          collection of cars, they are interested in exploring     the relationship between a set of variables and miles   per gallon (MPG) (outcome). They are particularly interested  in the following two questions:

 + “Is an automatic or manual transmission better for        MPG”

 + "Quantify the MPG difference between automatic and        manual transmissions"

# Load Library and Data

```{r, echo=TRUE, warning=FALSE, message=FALSE}
library(ggplot2)
data("mtcars")

head(mtcars)
```

# Exploratory Data analysis
```{r}
mtcars$am<-as.factor(mtcars$am)
levels(mtcars$am)<-c("AT", "MT")

aggregate(mpg~am, data=mtcars, mean)

ggplot(mtcars, aes(y = mpg, x = am, fill = am))+
      geom_boxplot()+
      labs( x = "Transmission Type", y = "Miles per Gallon", title = "")
```

# Statistical Inference
```{r}
auto_Data<-mtcars[mtcars$am == "AT",]
mannual_Data<-mtcars[mtcars$am == "MT",]
t.test(auto_Data$mpg, mannual_Data$mpg, paired = F, conf.level = 0.95)
```


# Regression Modelling

## Model fitting
```{r}
Reg_1 <- lm(mpg~am, mtcars)
summary(Reg_1)
```

```{r}
Reg_2 <- lm(mpg~am + wt, mtcars)
summary(Reg_2)
```

```{r}
Reg_3 <- lm(mpg~am+ wt+ factor(cyl), mtcars)
summary(Reg_3)
```

```{r}
Reg_All <- lm(mpg~., mtcars)

summary(Reg_All)
```

```{r}
par(mfrow = c(2,2))
plot(Reg_All)
```

# Conclusion 

### From Statistical inference
 + Since the p-value from the t-test was 0.0013, which is lower then alpha (5%). So We could reject the null hypothesis. which would imply that **there is a significant difference between manual and automatic transmission.** 

### From regresion model
 + From the simple regression model of am vs mpg(Reg_1), we could conclude that **manual cars have 7.24 MPG more than the automatic transmission car.**  
 with *R^2^* : 0.3598, the model only explains 35.98% of the variance. 



